---
title: "Homework 2"
author: "Jack Patnoe"
output:
    html_document:
      toc: true
      toc_float: true
      code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

Set Up: 



```{r, echo = FALSE, include = FALSE}

library(readr)
library(dplyr)
library(lmtest)
library(ggplot2)
library(gridExtra)
library(MASS)

```

Reading the Abalone data:

```{r, echo=TRUE}
abalone <- read.csv("/Users/jackpatnoe/Desktop/abalone.csv")

```

### Question 1

Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no `age` variable in the data set. Add `age` to the data set.

Assess and describe the distribution of `age`.

Adding age to our data set as # of rings + 1.5 : 

```{r, echo=FALSE}
abalone$age <- abalone$rings + 1.5
abalone %>%
  head()
hist(abalone$age, xlab = "Age", main = "Abalone Age Distr", col = "green" )
```
Looking at the Histogram for Abalone age, this appears to be a normal distribution with a slight right skew. The most common age is anywhere from 11 -12. 

### Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

```{r, echo=FALSE, include = FALSE}
library(tidyverse)
library(tidymodels)
```

Setting seed and splitting : 

```{r, echo=TRUE}
set.seed(1213)
abalone_split <- initial_split(abalone, prop = 0.80,
                                strata = age)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)
```



### Question 3

Using the **training** data, create a recipe predicting the outcome variable, `age`, with all other predictor variables. Note that you should not include `rings` to predict `age`. Explain why you shouldn't use `rings` to predict `age`.

Steps for your recipe:

1.  dummy code any categorical predictors

2.  create interactions between

    -   `type` and `shucked_weight`,
    -   `longest_shell` and `diameter`,
    -   `shucked_weight` and `shell_weight`

3.  center all predictors, and

4.  scale all predictors.

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.



```{r, echo = TRUE}
abalone_recipe <-
  recipe(age ~ type + shucked_weight + longest_shell + diameter + 
           shucked_weight + shell_weight, data = abalone) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_scale(all_predictors()) %>% step_center(all_predictors())
```

This is the recipe we can use to predict age. The reason we are not including rings is because it would kill the abalone, and we are trying to find a better way to predict age than doing that. 

### Question 4

Create and store a linear regression object using the `"lm"` engine.


```{r, echo = TRUE}
lm_model <- linear_reg() %>% 
  set_engine("lm")
```

### Question 5

Now:

1.  set up an empty workflow,
2.  add the model you created in Question 4, and
3.  add the recipe that you created in Question 3.

1. Settin up the empty workflow: 

```{r, echo = TRUE}
lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)
```

2. Using lm_flow from question 4 and our recipe:  

```{r, echo = TRUE}
lm_fit <- fit(lm_wflow, abalone_train)
```


```{r, echo = TRUE}
lm_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

### Question 6

Use your `fit()` object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

Using all of the given parameters we can make female age a new variable: 

```{r, echo = TRUE}
female_age <- data.frame(type = 'F', longest_shell = 0.50, diameter = 0.10,
                          height = 0.30, whole_weight = 4, shucked_weight = 1, 
                          viscera_weight = 2, shell_weight = 1)
predict(lm_fit, female_age)
```

Our prediction is that the age of this female abalone will be 16.39723, given these parameters we are given.  


### Question 7

Now you want to assess your model's performance. To do this, use the `yardstick` package:

1.  Create a metric set that includes *R^2^*, RMSE (root mean squared error), and MAE (mean absolute error).
2.  Use `predict()` and `bind_cols()` to create a tibble of your model's predicted values from the **training data** along with the actual observed ages (these are needed to assess your model's performance).
3.  Finally, apply your metric set to the tibble, report the results, and interpret the *R^2^* value.



```{r, echo = FALSE}
abalone_train %>%
  head()
```

```{r, echo = FALSE}
abalone_train_res <- predict(lm_fit, new_data = abalone_train)
abalone_train_res %>% 
  head()
```


```{r, echo = FALSE}
abalone_train_res <- bind_cols(abalone_train_res, abalone_train)
abalone_train_res %>% 
  head()
```
```{r, echo = FALSE}
rmse(abalone_train_res, truth = age, estimate = .pred)
abalone_metrics <- metric_set(rmse, rsq, mae)
abalone_metrics(abalone_train_res, truth = age, 
                estimate = .pred)
```
Assessing our models result, our RSQ value is .5207315, so this is a pretty good fit and our model did well. 
